---
title: "Billboard Graphic"
execute:
  echo: false
  warning: false
 
---

This graphic visualizes how the popularity of a selection of hit songs from the year 2000 evolved over time on the Billboard Hot 100 chart. Each line in the plot represents a song, showing its rank across the weeks it stayed on the chart. Lower values on the y-axis indicate better chart performance — for example, a rank of 1 means the song was number one that week. 

By analyzing this visualization, we can observe how quickly certain songs climbed the charts, how long they stayed at the top, and how their popularity changed week by week.

```{r}

suppressPackageStartupMessages(library(tidyverse))

```

```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
# Load the billboard dataset
data("billboard", package = "tidyr")

# View first few rows (optional)
#head(billboard)
# Pivot the data longer so each row is one week of ranking for a song
billboard_long <- billboard |>
  pivot_longer(cols = starts_with("wk"),
               names_to = "week",
               names_prefix = "wk",
               values_to = "rank",
               values_drop_na = TRUE) |>
  mutate(week = as.integer(week))  # Convert week to integer
# Select a few well-known songs
top_songs <- c("Smooth", "Breathe", "Say My Name", "Try Again", "Maria Maria")

# Plot
ggplot(filter(billboard_long, track %in% top_songs), 
       aes(x = week, y = rank, color = track)) +
  geom_line(linewidth = 1.2) +
  scale_y_reverse(breaks = seq(0, 100, 10)) +  # Billboard rank 1 is best
  scale_x_continuous(breaks = 1:20) +
  labs(title = "Billboard Chart Rankings Over Time (2000)",
       subtitle = "For a selection of top songs",
       x = "Week on Chart",
       y = "Rank (Lower is Better)",
       color = "Song") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```
